---
title: "Artical_sharing_EDA"
output: html_notebook
---

Deskdrop is an internal communications platform developed by CI&T, focused in companies using Google G Suite. Among other features, this platform allows companies employees to share relevant articles with their peers, and collaborate around them.

libraries used in the data set.
```{r}
library(ggplot2)
library(plotly)
library(data.table)
```

Reading the data into r
```{r}
users_interactions<-fread("users_intractions_format.csv")
shared_articles <- fread("shared_articles_format.csv")
```

converting the timestamp colum to date time column in r
```{r}
#strptime("2016-03-29 00:49:39",f)
users_interactions$timestamp<-as.POSIXct(strptime(users_interactions$timestamp, "%Y-%m-%d %H:%M:%S"))
shared_articles$timestamp<-as.POSIXct(strptime(shared_articles$timestamp,"%Y-%m-%d %H:%M:%S"))
```

```{r}
users_interactions[rev(order(timestamp)),]
```

```{r}
shared_articles[order(timestamp),]
```

EDA on user intraction

```{r}
length(unique(users_interactions$userRegion))
```
There 72 cities log data of user interactions.


```{r}
unique(users_interactions$userCountry)
```
Total we have 23 countries.

```{r}

```




