---
title: "The commute data of USA"
output: html_notebook
---

libraries
```{r}
library(data.table)
library(ggplot2)
library(plotly)
```

now reading the data into R 
```{r}
#census<-fread("",integer64="character")
commute<-fread("commute_data.csv") #integer64="character"
census<-fread("census_tracts_2010.csv")
```

Now viewing the 2 data sets.
```{r}
View(commute)
View(census)
```

```{r}
str(commute)
```
```{r}
str(census)
```
summary of the data set.
```{r}
summary(commute)
```
```{r}
summary(census)
```
combinig the 2 tables.
checking
```{r}
census[,.N,by=GEOID][,sum(N)]
```
Creating the final data.

now we are joining the two tables.
now we are re-naming the column from senser table
```{r}
colnames(census)[colnames(census) == 'GEOID'] <- 'OFIPS'
```

doing inner join based on OFIPS column in commute data.
```{r}
d1<-merge(commute,census, by = "OFIPS")
```

again renaming the column to 
```{r}
colnames(census)[colnames(census) == 'OFIPS'] <- 'DFIPS'
```

now creating d2
```{r}
d2<-merge(commute,census, by = "DFIPS")
```

The two data frames are not identical.
```{r}
identical(d1, d2)
```
Now creating the final Data.




```{r}
d1<-d1[order(d1$OFIPS)]
```

```{r}
d2<-d2[order(d2$OFIPS)]
```

```{r}
colnames(d2)
```
changing the column order.
```{r}
col_order<-c(   "OFIPS" ,  "DFIPS"   , "OSTFIPS"    , "OCTFIPS"  ,   "OTRFIPS"  ,   "DSTFIPS" ,    "DCTFIPS", "DTRFIPS"  ,   "FLOW"   ,     "MOE"    ,     "LENKM"    ,   "ESTDIVMOE" ,  "USPS"     ,   "POP10"  ,    
 "HU10"     ,   "ALAND"     ,  "AWATER"    ,  "ALAND_SQMI" , "AWATER_SQMI" ,"INTPTLAT"   , "INTPTLONG" )
d2<-as.data.frame(d2)
d2 <- d2[,col_order]
d2<-as.data.table(d2)
```

intersection.
```{r}
#d3<-fintersect(setDT(d1), setDT(d2),all = FALSE)
```

union.
```{r}
final<- funion(d1,d2)
```









Now on this data we are going to do our final analysis.
```{r}
str(final)
```
```{r}
sum(is.na(final))
```
```{r}

```

