---
title: "BART Ridership Analysis"
output: html_notebook
---

# About domain:

BART, short for "Bay Area Rapid Transit", is the transit system severing the San Francisco Bay Area in California. BART operates six routes, 46 stations, and 112 miles of track. It serves an average weekday ridership of 423,000 people, making it the fifth-busiest rapid transit system in the United States.

# Problem 

The BART organization wants to know some insights from the data of 2016 till May of 2017, some insights like:-

1) Which BART station is the busiest?
2) When is the best time to go to SF if you want to find a seat?
3) Which day of the week is the busiest?
4) How many people take the BART late at night?  etc......

# Solution provided/Opportunity

By using Analytics we can get insights from the data and by these insights, the organization can better allocate their resources.we are also forecasting weekly average of ridership.


# Technologies Used :

we have used Data science to forecast average weekly Ridership.


# About the data set

Note: - There are 2 data sets used in this analysis.

#### Data set 1 :- 

This dataset contains daily information on BART ridership between stations for a period covering all of 2016 to till May 2017. this data includes movements between specific stations (there are just over 2000 station-to-station combinations).

#### Data set 2 :-

This dataset contains information on ridership weekly average from July 1997 to May  2018

#### Source of data set

  1) Data set 1 :-  https://www.kaggle.com/saulfuh/bart-ridership/data.

  2) Data set 2 :- https://data.bart.gov/dataset/customer-ridership/resource/6e653520-58cf-45c5-b40c-d37c8957ec77.

#Process/steps followed

  step 1:- Plotted over insights of BART stations

  step 2 :- Plotted route wise insights.

  step 3 :- Forecasting weekly Ridership till Dec 2019.

#### Models used

 ARIMA.

# Over all Inferences

Note:- The Origin station and destination station are given by their respective codes in the plotted graphs.
```{r,echo=FALSE}
library(data.table)
library(ggplot2)
library(plotly)
library(zoo)
library(lubridate)
library(dplyr)
library(quantmod)
library(reshape2)
library(plyr)
library(scales)
library(RColorBrewer)
#library(RColorBrewer)
library(devtools)
library(scales)
library(tseries)
library(forecast)
#for intractive plots of time series.
library(dygraphs)
library(xts)
library(zoo)
library(dplyr)
library(tidyr)
```

```{r,echo=FALSE}
season_plot<-function(start,end){
data[Origin==start & Destination==end & year==2016,list(Throughput=sum(Throughput)),by=c("season")] %>%
      group_by(season)  %>%
      plot_ly(labels = ~as.factor(season), values = ~Throughput) %>%
      add_pie(hole = 0.6) %>%
      layout(title = "Percentage passengers on different seasons of year 2016",  showlegend = T,
             xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
}
```

```{r,echo=FALSE}
time_interval_function<-function(start,end){
data[Origin==start & Destination==end,list(Throughput=sum(Throughput)),by=c("time_intervals_d")][order(Throughput,time_intervals_d,decreasing=TRUE)]  %>%group_by(time_intervals_d)  %>%
      plot_ly(labels = ~as.factor(time_intervals_d), values = ~Throughput) %>%
      add_pie(hole = 0.6) %>%
      layout(title = "Percentage of passengers at different time intervals",  showlegend = T,
             xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
}
```



```{r,echo=FALSE}
fte_theme <- function() {
  
  # Generate the colors for the chart procedurally with RColorBrewer
  palette <- brewer.pal("Greys", n=9)
  color.background = palette[2]
  color.grid.major = palette[3]
  color.axis.text = palette[6]
  color.axis.title = palette[7]
  color.title = palette[9]
  
  # Begin construction of chart
  theme_bw(base_size=15) +
    
    # Set the entire chart region to a light gray color
    theme(panel.background=element_rect(fill=color.background, color=color.background)) +
    theme(plot.background=element_rect(fill=color.background, color=color.background)) +
    theme(panel.border=element_rect(color=color.background)) +
    
    # Format the grid
    theme(panel.grid.major=element_line(color=color.grid.major, size=.05)) +
    theme(panel.grid.minor=element_blank()) +
    theme(axis.ticks=element_blank()) +
    
    # Format the legend, but hide by default
    theme(legend.background = element_rect(fill=color.background, size=.5, linetype="dotted")) +
    theme(legend.text = element_text(size=8,color="black")) +
    theme(legend.position = "top") +
    theme(legend.title=element_blank()) +
   
    
    # Set title and axis labels, and format these and tick marks
    theme(plot.title=element_text(color=color.title, size=20, vjust=4.25, face = "bold")) +
    theme(axis.text.x=element_text(size=14,color="black")) +
    theme(axis.text.y=element_text(size=12,color="black")) +
    theme(axis.title.x=element_text(size=14,color="black", face = "bold")) +
    theme(axis.title.y=element_text(size=14,color="black", vjust=1.25)) +
    
    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}
```

```{r, echo=FALSE}
season_graph<-function(route){
ggplot(data[routes==route,list(Throughput=sum(Throughput)),by=c("season")], aes(as.factor(season),Throughput)) + geom_tile(aes(fill = Throughput),colour = "red") + scale_fill_gradient(low = "yellow", high = "red")+labs(title = paste("The Trend of Throughput on routes by seasons",root))+xlab("Seasons")+ylab("Throughput")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "Black"))
}
```

```{r, echo=FALSE}
clender_map<-function(origin,destination){
ggplot(data[Origin==origin & Destination==destination,list(Throughput=sum(Throughput)),by=c("monthweek","weekDays","monthf","year")], aes(monthweek,factor(weekDays,levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday" )), fill = Throughput)) + 
  geom_tile(colour = "white") + facet_grid(year~factor(monthf,levels = c("Jan", "Feb" ,"Mar" ,"Apr" ,"May" ,"Jun" ,"Jul" ,"Aug" ,"Sep" ,"Oct" ,"Nov" ,"Dec"))) + scale_fill_gradient(low="yellow", high="red") +
  ggtitle(paste("The Busiest time from",origin,"to",destination))+labs(fill="No of passengers")+xlab("\n\nWeek of Month")+ ylab("")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "black"))
}
```

```{r, echo=FALSE}
monthweeks <- function(x) {
    UseMethod("monthweeks")
}
monthweeks.Date <- function(x) {
    ceiling(as.numeric(format(x, "%d")) / 7)
}
monthweeks.POSIXlt <- function(x) {
    ceiling(as.numeric(format(x, "%d")) / 7)
}
monthweeks.character <- function(x) {
    ceiling(as.numeric(format(as.Date(x), "%d")) / 7)
}
```

```{r echo=FALSE}
trendGraph<-function(route){
ggplotly(ggplot(data[routes==route,list(Throughput=sum(Throughput)),by=c("month","year")],aes(x=month,y=Throughput,group=1))+ geom_line(linetype = "dashed",color="red")+labs(title = paste("The Trend of Passengers on route",route))+xlab("Month")+ylab("No Passengers")+geom_point()+facet_grid(. ~ year)+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "black"))+theme(axis.title.x = element_text(vjust=-0.1)))
}
```

```{r, echo=FALSE}
data1 <- fread("date-hour-soo-dest-2016.csv",stringsAsFactors = TRUE)
data2 <- fread("date-hour-soo-dest-2017.csv",stringsAsFactors = TRUE)
data3 <- fread("station_info.csv",stringsAsFactors = TRUE)
```

```{r, echo=FALSE}
data<-rbind(data1,data2)
rm(data1)
rm(data2)
```

```{r ,echo=FALSE}
data$DateTime  <- as.POSIXct(data$DateTime, format="%Y-%m-%d %H:%M:%S")
```

```{r echo=FALSE}
data$year=year(data$DateTime)
data$month=as.factor(month(data$DateTime))
data$weekDays=weekdays(data$DateTime)
data$hours = hour(data$DateTime)
```

```{r ,echo=FALSE}
indx <- setNames( rep(c('winter', 'spring', 'summer',
                   'fall'),each=3), c(12,1:11))
#data1=data
data$season<-unname(indx[as.character(data$month)])
```

```{r ,echo=FALSE}
# create breaks
breaks <- hour(hms("00:00", "6:00", "12:00", "18:00", "23:59"))
# labels for the breaks
labels <- c("12AM-6AM", "6AM-12Noon", "12Noon-6pm", "6pm-12Am")
data$time_intervals <- as.factor(cut(x=hour(data$DateTime), breaks = breaks, labels = labels, include.lowest=TRUE))
```

```{r,echo=FALSE}
#data$Destination<-as.factor(data$Destination)
#data$Origin<-as.factor(data$Destination)
```

```{r ,echo=FALSE}
data1=data[,list(Throughput=sum(Throughput)), by='Destination'][order(Throughput,Destination,decreasing=TRUE)]
```




```{r ,echo=FALSE}
x <-ggplot(data1[1:5,],aes(x=Destination,y=Throughput))+
  geom_segment( aes(x=Destination, xend=Destination, y=0, yend=Throughput)) +ylim(0,22000000)+scale_y_continuous(labels = comma)+
  geom_point( size=5, color="steelblue", fill=alpha("steelblue", 0.3), alpha=0.7, shape=21, stroke=2)+theme(panel.background = element_rect(fill = ' white'))+geom_text(aes(label=Throughput), vjust=-2, color="black", size=3.4)+coord_flip()+labs(title = "The Top 5 populated BART Stations",x="Station code",y="Population")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "Black"))
```

```{r ,echo=FALSE}
ggsave("top_5_stations.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
***Legend***

EMBR :- Embarcadero (EMBR).

Mont :- Montgomery St. (MONT).

Powl :- Powell St. (POWL).

Civc :- Civic Center/UN Plaza (CIVC).

DBRK :- Downtown Berkeley (DBRK).

***Inference*** :-  As we see Embarcadero (EMBR) is the BART station was the busiest station in 2016 till may 2017, followed by Montgomery St. (MONT), Powell St. (POWL), Civic Center/UN Plaza (CIVC) and Downtown Berkeley (DBRK).
```{r ,echo=FALSE}
x<-ggplot(data1[41:45,],aes(x=as.factor(Destination),y=Throughput))+
  geom_segment( aes(x=as.factor(Destination), xend=as.factor(Destination), y=0, yend=Throughput)) +ylim(0,2200000)+
  geom_point( size=5, color="steelblue", fill=alpha("steelblue", 0.3), alpha=0.7, shape=21, stroke=2)+theme(panel.background = element_rect(fill = ' white'))+geom_text(aes(label=Throughput), vjust=-1, color="black", size=3.4)+coord_flip()+labs(title = "The Least 5 populated BART Stations",x="Station code",y="Population")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "Black"))
```

```{r ,echo=FALSE}
ggsave("Least_5_stations.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
***Legend***

SHAY - South Hayward (SHAY).

ORIN - Orinda (ORIN).

OAKL - Oakland Airport (OAKL).

NCON - North Concord/Martinez (NCON).

CAST - Castro Valley (CAST).


***Inference*** :-  The least  stations are Oakland Airport (OAKL).
```{r ,echo=FALSE}
rm(data1)
```


```{r ,echo=FALSE}
data$routes<-as.factor(with(data, paste0(Origin,"-",Destination)))
```

```{r ,echo=FALSE}
x<-ggplot(data[!(Origin==Destination),list(Throughput=sum(Throughput)),by=c("routes")][order(Throughput,routes,decreasing=TRUE)][1:5,],aes(x=routes,y=Throughput))+geom_bar(fill="steelblue",bin = 50,width = 0.2 ,stat="identity",na.rm = TRUE)+scale_y_continuous(labels = comma)  +geom_text(aes(label=Throughput), vjust=-1, color="black", size=3.4)+labs(title = "The top 5 busy routes of BART",x="Routes",y="Populations")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "black"))+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+coord_flip()
```

```{r,echo=FALSE}
ggsave("top_5_routes.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
***Legend***

POWL - BALB :- Powell St. (POWL) to Balboa Park (BALB).

POWL - 24th :- Powell St. (POWL) to 24th St. Mission (24TH).

DUBL - EMBR :- Dublin/Pleasanton (DUBL) to Embarcadero (EMBR).

BALB - POWL :- Balboa Park (BALB) to Powell St. (POWL)

***Inference*** :- Powell St. (POWL) to Balboa Park (BALB) is the busiest route, 1105884 passengers have traveled in the year 2016 till May 2017.
```{r ,echo=FALSE}
x<-ggplot(data[!(Origin==Destination),list(Throughput=sum(Throughput)),by=c("routes")][order(Throughput,routes,decreasing=TRUE)][2065:2070,],aes(x=routes,y=Throughput))+geom_bar(fill="steelblue",bin = 50,width = 0.2 ,stat="identity",na.rm = TRUE)+geom_text(aes(label=Throughput), vjust=-1.4, color="black", size=3.4)+labs(title = "Least popular Routes",x="Routes",y="Populations")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "black"))+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+coord_flip()

```

```{r ,echo=FALSE}
ggsave("least_5_routes.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
***Legend***

WSPR-SSAN :- Warm Springs/South Fremont (WARM) to South San Francisco (SSAN).

WSPR-SBRN :- Warm Springs/South Fremont (WARM) to  San Bruno (SBRN).

WSPR-MLBR :- Warm Springs/South Fremont (WARM) to Millbrae (MLBR).

SSAN - WSPR :- South San Francisco (SSAN) to Warm Springs/South Fremont (WARM).

SBRN - WSPR :- San Bruno (SBRN) to Warm Springs/South Fremont (WARM).

MLBR - WSPR :- Millbrae (MLBR) to Warm Springs/South Fremont (WARM).



***Inference***  :- These are the least 6 popular routes. Only 40 to 50 passengers have travelled in these routes in the time period of 2016 to  May of 2017.



```{r ,echo=FALSE}
x<-data[,list(Throughput=sum(Throughput)),by=c("time_intervals")][order(Throughput,time_intervals,decreasing=TRUE)]  %>%
  group_by(time_intervals)  %>%
  plot_ly(labels = ~time_intervals, values = ~Throughput) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Percentage of passengers on TimeIntervals of the day",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```


```{r}
x
#orca(x, file = "percentage_time.png")
#htmlwidgets::saveWidget(x, file = "percentage_time.html")
```


***Inference*** :- In the year 2016 till May 2017 41.3% of passengers have traveled at the time range of 12Noon-6pm and 37.2% of passengers have traveled at 6 am to 12 Noon.
```{r ,echo=FALSE}
data$monthf<-factor(data$month,levels=as.character(1:12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE)
```

```{r ,echo=FALSE}
data$yearmonth <- as.yearmon(data$DateTime)
data$yearmonthf <- factor(data$yearmonth)
data$week <- as.numeric(format(data$DateTime,"%W"))
data$monthweek<-monthweeks.POSIXlt(data$DateTime)
data<-as.data.table(data)
```

Note:- These are the following assumptions for the below graph

1) Spring:- March, April, May.

2) Summer:- June, July, August.

3) Autum:- September, October, November.

4) Winter:- December, January, February.


```{r ,echo=FALSE}
x<-data[year==2016,list(Throughput=sum(Throughput)),by=c("season")] %>%
  group_by(season)  %>%
  plot_ly(labels = ~as.factor(season), values = ~Throughput) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Percentage of passengers on different seasons on 2016",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
x
#htmlwidgets::saveWidget(x, file = "percentage_seasons.html")
```
***Inference*** :- In 2016 there is very less variation in number of passengers on different seasons.

```{r ,echo=FALSE}
x<-ggplot(data[,list(Throughput=sum(Throughput)),by=c("monthweek","weekDays","monthf","year")], aes(monthweek,factor(weekDays,levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday" )), fill = Throughput)) + 
  geom_tile(colour = "white") + facet_grid(year~factor(monthf,levels = c("Jan", "Feb" ,"Mar" ,"Apr" ,"May" ,"Jun" ,"Jul" ,"Aug" ,"Sep" ,"Oct" ,"Nov" ,"Dec")))+ scale_fill_gradient(low="yellow", high="red")+scale_fill_continuous(labels = comma) +
  ggtitle(paste("The Busiest time"))+labs(fill="No of passengers")+xlab("\n\nWeek of Month")+ ylab("")+theme(plot.title = element_text(hjust = 0.5,face = "bold",color = "black"))
```
```{r ,echo=FALSE}
#ggsave("heat_map1.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
The above chart shows the Number of passengers traveled by months, weekdays, week, month and year.Light blue represents the passengers are high, dark blue represents the passengers are low.

***Inference*** :- In 2016 till may of 2017 generally, the people have used BART more in weekdays as compared to weekends.so if you want to find a seat for yourself, the best time is to travel at weekends.


# Insights by routes.

```{r,echo=FALSE}
# create breaks
breaks <- hour(hms("00:00","3:00", "6:00","9:00","12:00","15:00","18:00","21:00","23:59"))
# labels for the breaks
labels <- c("12Am-3Am","3Am-6Am", "6Am-9Am","9Am-12Noon","12Noon-3Pm","3Pm-6Pm","6Pm-9Pm","9Pm-12Am")
data$time_intervals_d <- as.factor(cut(x=hour(data$DateTime), breaks = breaks, labels = labels, include.lowest=TRUE))
```

Here we have done analysis by route, so here analysis have done on  Powell St. (POWL) to Balboa Park (BALB) as it is busiest route as compare to other routes.

#### Powell St. (POWL) to Balboa Park (BALB) 
```{r ,echo=FALSE}
x<-time_interval_function("POWL","BALB")
#htmlwidgets::saveWidget(x, file = "percentage_tim.html")
x
```
***Inference*** :- In this route, 38% of passengers have traveled at the time interval of 3pm-6pm and 19% of people have traveled at time interval 6Pm to 9pm.
```{r ,echo=FALSE}
data$DateTime<-data[order(data$DateTime),]$DateTime
data<-as.data.table(data)
```

```{r ,echo=FALSE}
x<-trendGraph("POWL-BALB")
htmlwidgets::saveWidget(x, file = "trend.html")
x
```
***Inference*** :-In between the year the trend for this route is decreasing, At the end of the years the trend for this route increasing.
```{r ,echo=FALSE}
x<-clender_map("POWL","BALB")
#ggsave("heat_map2.jpg",plot=x,unit="in",width=7,height = 7,dpi=300)
x
```
***Inference*** :- In this route, the people are traveling more on weekdays as compared to weekends.so if you want to find a seat for yourself, the best time is to travel weekends of August and september.


# Forecasting of weekly Ridership

Note:- we have used Data set 2 for the forecasting.
```{r ,echo=FALSE}
data<-fread("customer-ridership.csv")
```

```{r ,echo=FALSE}
data$FiscalYearMonth <- paste(data$FiscalYear,data$FiscalMonth, sep="")
data$FiscalYearMonth<-as.yearmon(as.character(data$FiscalYearMonth),"%Y%m")
```

```{r ,echo=FALSE}
data_time_series<-xts(x = data$`RIDERSHIP WEEKAVG` ,order.by = data$FiscalYearMonth)
```

```{r ,echo=FALSE}
dygraph(data_time_series,main = "The Trend of average weekly Ridership") %>%
      dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors="#D8AE5A") %>%
      dyRangeSelector() %>%
      dyCrosshair(direction = "vertical") %>%
      dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
      dyRoller(rollPeriod = 1)%>%dyAxis("y", label = "Ridership")%>%dyAxis("x",label = "Years")
```
***Inference*** :- As we see the trend of weekly average ridership went on increasing from 1997 to June of 2018.


```{r ,echo=FALSE}
#converting the data into log form to make variation constant.
data_time_series<-log(data_time_series)
    
#getting data into proper format.
data_time_series<- ts(data_time_series, frequency = 12, start = c(1997,1))
    
#fitting the model
model<-auto.arima(data_time_series)
    
#forcasting the value
forcast_values<-forecast(model,h=17)
    
#Extracting the forecasted values
forcast_values_extracted<-as.numeric(forcast_values$mean)

#here we are converting the log values into normal form.
Forecasted_Values<-exp(forcast_values_extracted)
    
#converting the values to dataFrame
Forecasted_Values<-as.data.frame(Forecasted_Values)

```

By using ARIMA Model we have forecasted weekly average of Ridership from August 2018 to December 2019.

These are the forecasted values of weekly average of Ridership.

```{r ,echo=FALSE}
round(Forecasted_Values)
```

**Forecasted_Values of average weekly ridership.**

August 2018 :-  414608

september 2018 :- 404722

october 2018 :- 404229

November 2018 :- 407280

December 2018 :- 407841

January 2019 :- 404122

February 2019 :- 403776

March 2019 :- 409430

Aprial 2019 :- 410061

May 2019 :- 407389

June 2019 :- 386703

July 2019 :-  387868

August 2019 :- 400749

september 2019 :- 396350

october 2019 :- 394525	

November 2019 :- 397764	

December 2019 :- 398429	
```{r ,echo=FALSE}
    
#getting data into proper format.
data_time_series<- ts(data_time_series, frequency = 12, start = c(1997,1))
    
#fitting the model
model<-auto.arima(data_time_series)
    
#forcasting the value
forcast_values<-forecast(model,h=17)
    
#preparing the data for graphs
actuals <- forcast_values$x
lower <- forcast_values$lower[,2]
upper <- forcast_values$upper[,2]
point_forecast <- forcast_values$mean
    
f<-cbind(actuals, lower, upper, point_forecast)
    
```
#### Note :- The prediction value are plotted on graph in log forms.
```{r,echo=FALSE}
dygraph(f, main = "The forecast of weekly average ridership") %>% 
dyRangeSelector() %>% 
dyRangeSelector(height = 40,
dateWindow = c("2015-04-01", "2021-4-01")) %>%
dySeries(name = "actuals", label = "actual") %>%
dySeries(c("lower","point_forecast","upper"), label = "Predicted") %>%
dyLegend(show = "always", hideOnMouseOut = FALSE) %>%
dyHighlight(highlightCircleSize = 5,
highlightSeriesOpts = list(strokeWidth = 2)) %>%
dyOptions(axisLineColor = "navy", gridLineColor = "grey")%>%dyAxis("y", label = "Ridership(Values are in log form)")%>%dyAxis("x",label = "Years")
```

We have used ARIMA and Holt winter for forecasting, on which ARIMA gave better results for our data with the error deviation of 4%.
 
# Conclusion

We have properly analyzed the data and extracted the insights by which BART organization can better plan their resources ,we have also successfully build forecasting mode and forecasted average weekly ridership from August 2018 to till july 2020.  


