---
title: "Leaf Let maps tutorials"
output: html_notebook
---

#Leaf Let Tutorial.

```{r}
#install.packages('leaflet')
library(leaflet)
```

```{r}
leaflet() %>% addTiles() %>% addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```
#Initializing Options

The map widget can be initialized with certain parameters. This is achieved by populating the options argument as shown below.
```{r}
# Set value for the minZoom and maxZoom settings.
leaflet(options = leafletOptions(minZoom = 0, maxZoom = 18))
```

```{r}
# add some circles to a map
df = data.frame(Lat = 1:10, Long = rnorm(10))
leaflet(df) %>% addCircles()
```
```{r}
leaflet(df) %>% addCircles(lng = ~Long, lat = ~Lat)
```
```{r}
leaflet() %>% addCircles(data = df)
leaflet() %>% addCircles(data = df, lat = ~ Lat, lng = ~ Long)
```

```{r}
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```

```{r}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```

```{r}
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)
m %>% addTiles()
```

```{r}
m %>% addProviderTiles(providers$Stamen.Toner)
```

```{r}
m %>% addProviderTiles(providers$CartoDB.Positron)
```

```{r}
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```
```{r}
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 4) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```
Combining Tile Layers
```{r}
m %>% addProviderTiles(providers$MtbMap) %>%
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stamen.TonerLabels)
```

```{r}
data(quakes)

# Show first 20 rows from the `quakes` dataset
leaflet(data = quakes[1:20,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```
#Customizing Marker Icons

You can provide custom markers in one of several ways, depending on the scenario. For each of these ways, the icon can be provided as either a URL or as a file path.

For the simple case of applying a single icon to a set of markers, use makeIcon().
```{r}
greenLeafIcon <- makeIcon(
  iconUrl = "http://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

```{r}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "http://leafletjs.com/examples/custom-icons/leaf-green.png",
    "http://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = leafIcons)
```

```{r}
# Make a list of icons. We'll index into it based on name.
oceanIcons <- iconList(
  ship = makeIcon("ferry-18.png", "ferry-18@2x.png", 18, 18),
  pirate = makeIcon("danger-24.png", "danger-24@2x.png", 24, 24)
)

# Some fake data
df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 90.620130,  # lng
    (runif(20) - .5) * 3.8 + 25.638077  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "pirate", "ship"),
    c("ship", "pirate")
  ))
)

leaflet(df) %>% addTiles() %>%
  # Select from oceanIcons based on df$type
  addMarkers(icon = ~oceanIcons[type])
```
```{r}
# first 20 quakes
df.20 <- quakes[1:20,]

getColor <- function(quakes) {
  sapply(quakes$mag, function(mag) {
  if(mag <= 4) {
    "green"
  } else if(mag <= 5) {
    "orange"
  } else {
    "red"
  } })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(df.20)
)

leaflet(df.20) %>% addTiles() %>%
  addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(mag))
```

#Cluster options
```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions()
)
```

```{r}
leaflet(df) %>% addTiles() %>% addCircleMarkers()
```

```{r}
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )
```

#Popups

```{r}

```

