---
title: "Data Cleaning on weather data set."
output: html_notebook
---



#About the data set.

It's covers hourly weather data from 122 weathers stations of southeast region (Brazil). The southeast include the states of Rio de Janeiro, São Paulo, Minas Gerais e Espirito Santo.


libraries needed
```{r}
#used for reading the data ,data subsetting
library(data.table)
#used for anabling parallel processing
library(doMC)
library(rapportools)
```

As the data set is large so we are doing parallel processing.
```{r}
registerDoMC(cores = 6)
```


so now we are importing the into R.
As the data is huge the due to limitations of hardware support we are considering our data from 2012 to 2016
```{r}
data<-fread("sudeste.csv")
#subsetting the data 
#we are subsetting our data.
data<-data[yr>2011]
```

```{r}
cat("\nThe Tota number of row in the data set :- ",nrow(data))
cat("\nThe Total number of columns in the data set :- ",ncol(data))
```
so as we see above that total number of row we have is 4745736,and the column we have is 31.

Now we are going to look at summary of the data set.
```{r}
summary(data)
```
Just viewing the columns for more clarity
```{r}
colnames(data)
```
checking whether there are any blank spaces in the data set.
```{r}
#Checking the NA 'S at entire data set.
cat("The Number of NA 'S in the data: ", sum(data=="NA"))
#checking for the empty sells in the entire data set.
cat("\nThe Number of blank cell 'S in the data: ", sum(data==""))
```
```{r}
apply(data,2,function(x)sum(is.na(x)))
```
```{r}
apply(data,2,function(x)sum(data==""))
```

```{r}
#This command is use to see all the empty rows.
data[gbrd==""]
```
```{r}
data[is.na(prcp)==TRUE]
```
```{r}
#is.empty(data$temp)
data[is.empty(prcp)==TRUE]
```

```{r}
rowSums()
```

