    ---
title: "R Notebook"
output: html_notebook
---

```{r}
library(data.table)
library(stringr)
library(dplyr)
library(leaflet)
# library(readxl)
```
Importing the data
```{r}
# sample_data <- read_excel("sample_data.xlsx")
tran <- fread("sample_data.csv",na.strings=c("NA", "null",""))
vehicles_info <- as.data.frame(tran$VehiclePosition)
```

Trying to read the vehicle position data in detail
```{r}
vehicles_info <- as.data.frame(str_split_fixed(vehicles_info$`tran$VehiclePosition`, ",", 18))

```

Treating the last column and making anew dataframe
```{r}
x <- lapply(split(unlist(
  lapply(vehicles_info,tstrsplit,'":'),recursive=FALSE),c("Left","Right")),
  as.data.frame)
data_1 <- x$Left


temp <- as.data.frame(gsub('desi:"' ,"",data_1$V1))

data_2 <- cbind(temp,data_1[,-1])
data_3 <- data.frame(lapply(data_2, function(x) {
                  gsub('"', "", x)
              }))

col_name <- x$Right[1,]

col_name <- lapply(col_name, function(x) {
                  gsub('"', "", x)})

 
```

Giving col names to th new formed data frame
```{r}

col_name <- c("desi", col_name)
colnames(data_3) <- col_name

```

Understanding the original dataset 

```{r}
tran_1 <- tran[, -which(names(tran) %in% "VehiclePosition")]#removing the last coln

str(tran_1)
colSums(is.na(tran_1))
```

```{r}

apply(tran_1, 2, function(x) length(unique(x)))

```
Finding the unique value
```{r}
apply(tran_temp, 2, function(x) unique(x))
```

Exploring the payload data
```{r}

apply(data_3, 2, function(x) length(unique(x)))
colSums(is.na(data_3))

```

Comparing original dataset and payload dataset for various variables
```{r}
sum(ifelse(tran_1$operator_id==data_3$oper,0,1)) #operator ID
sum(ifelse(tran_1$vehicle_number==data_3$veh,0,1)) #Vehicle number
sum(ifelse(tran_1$direction_id==data_3$dir,0,1)) #direction ID
sum(ifelse(tran_1$start_time==data_3$start,0,1)) #start time
sum(ifelse(tran_1$temporal_type==data_3$dir,0,1)) #direction

```
Considering the GTFS route_id for mapping the route 
```{r}
imp_original <- tran[,which(names(tran) %in% c("transport_mode","vehicle_number",
                                               "route_id","direction_id","next_stop",
                                               "geohash_level","geohash"))]

imp_payload <- data_3[,which(names(data_3) %in% c)]

data_map <- cbind()
       
```

```{r}
str(data_3)
```
subsetting the important features
```{r}
data_main1<-data_3[,c("tst","lat","long","start")]
data_main2<-tran[,c("temporal_type","transport_mode","operator_id" , "vehicle_number" ,"route_id" ,"start_time", "next_stop")]
```

now plotting the leaflet map.
```{r}
View(data_main1)
View(data_main2)
```

combining the 2 tables.
```{r}
data_main<-cbind(data_main1,data_main2)
```

checking the na values in the table
```{r}
sum(is.na(data_main))
```
checking the structure and summary of the data
```{r}
str(data_main)
summary(data_main)
```

null value treatment
```{r}

which(data_main$lat  == "NA")
#na_data<-data_main[which(data_main$lat=="null" | data_main$long=="null"),]
#na_data<-data_main[which(data_main == "NA" | data_main == "null" | data_main ==""),]
```
type casting the variables
```{r}
data_main[,c("lat","long")]<-lapply(data_main[,c("lat","long")],as.character)
data_main[,c("lat","long")]<-lapply(data_main[,c("lat","long")],as.numeric)
```

```{r}
str(data_main)
summary(data_main)
```









shiny part
```{r}
ui<-

```




