---
title: "ALL operations by Data.table"
output: html_notebook
---
libraries
```{r}
library("data.table")
library("curl")
```
Now reading the data into R session
```{r}
flights<-fread("https://raw.githubusercontent.com/Rdatatable/data.table/master/vignettes/flights14.csv",header=TRUE,sep=",",na.strings = c('?','-','*','$','^','NA'))
```
creating user define data-frame by data.table.
```{r}
#one way of creating the dataframe is this :-
DT = data.table(
  ID = c("b","b","b","a","a","c"),
  a = 1:6,
  b = 7:12,
  c = 13:18
)
#The other way of creating the data frame is this :-
ID = c("b","b","b","a","a","c")
a = 1:6
b = 7:12
c = 13:18
DT1=data.table(ID,a,b,c)
DT1
```
The 
```{r}
str(flights)
View(flights)
```

```{r}
summary(flights)
```
The general format of the data.table :- 


subsetting the row in the dataframe
```{r}
ans <- flights[origin == "JFK" & month == 6L,]
head(ans)
```
getting first two rows from the data frame.
```{r}
ans <- flights[1:2]
ans
```
sorting the dataframe by the column or columns
- :- des
+ :- asc
THE ORDER FUNCTION :- 

The  function order the first parameter first,by refering the class of first parameter it order the scond parameter.
example :-
param1     param2
A           Z
A           Y 
A           X
A           W
A           V
A           U
B           Z
B           Y
B           X
```{r}
ans_des <- flights[order(origin, -dest)]
ans_asc <- flights[order(origin, dest)]
head(ans_des)
head(ans_asc)
```
Selecting the columns.
this returns a vector.
```{r}
ans <- flights[, arr_delay]
ans
```
This will return the dataframe.
```{r}
ans <- flights[, list(arr_delay)]
print(class(ans))
ans
```
This method will also return the dataframe ,(suraj sir said not use this method)(Reason :- for code readability)
```{r}
ans <- flights[,.(arr_delay)]
print(class(ans))
ans
```
Renaming the column.
```{r}
ans <- flights[,list(delay_arr = arr_delay, delay_dep = dep_delay)]
ans
```
How many trips have had total delay < 0?
```{r}
ans <- flights[, sum( (arr_delay + dep_delay) < 0 )]
ans
```
Subset in i and do in j
```{r}
ans <- flights[origin == "JFK" & month == 6L,
               list(m_arr = mean(arr_delay), m_dep = mean(dep_delay))]
ans
```
How many trips have been made in 2014 from "JFK" airport in the month of June?
```{r}
ans <- flights[origin == "JFK" & month == 6L, length(dest)]
ans
```
using the special symbols getting the above example.
(.N) is the special symbol 
.N is a special built-in variable that holds the number of observations in the current group.
```{r}
ans <- flights[origin == "JFK" & month == 6L, list(.N)]
ans
```
Great! But how can I refer to columns by names in j (like in a data.frame)?
```{r}
ans <- flights[, c("arr_delay", "dep_delay")]
head(ans)
```
- Select columns named in a variable using the .. prefix
```{r}
select_cols = c("arr_delay", "dep_delay")
flights[ , ..select_cols]
```
Select columns named in a variable using with = FALSE
```{r}
flights[ , select_cols, with = FALSE]
```
using with function in "dataframe"
Hence the argument name with in data.table. Setting with = FALSE disables the ability to refer to columns as if they are variables, thereby restoring the "data.frame mode".
```{r}
DF = data.frame(x = c(1,1,1,2,2,3,3,3), y = 1:8)
DF
DF[with(DF, x > 1), ]
```
We can also deselect columns using - or !. For example:
```{r}
# returns all columns except arr_delay and dep_delay
ans <- flights[, !c("arr_delay", "dep_delay")]
# or
ans <- flights[, -c("arr_delay", "dep_delay")]
ans
```
we can also select by specifying start and end column names, e.g., year:day to select the first three columns.
```{r}
## not run

# returns year,month and day
ans <- flights[, year:day]
ans
# returns day, month and year
ans <- flights[, day:year]
ans
# returns all columns except year, month and day
ans <- flights[, -(year:day)]
ans
ans <- flights[, !(year:day)]
ans
```
2. Aggregations
a) Grouping using by
- How can we get the number of trips corresponding to each origin airport?
```{r}
ans <- flights[, list(.N), by = list(origin)]
ans
```

```{r}
ans <- flights[, .N, by = origin]
ans
```
- How can we calculate the number of trips for each origin airport for carrier code "AA"
```{r}
ans <- flights[carrier == "AA", .N, by = origin]
ans
```
How can we get the total number of trips for each origin, dest pair for carrier code "AA"?
```{r}
ans <- flights[carrier == "AA", .N, by = list(origin, dest)]
head(ans)
```
 How can we get the average arrival and departure delay for each orig,dest pair for each month for carrier code "AA"?
```{r}
ans <- flights[carrier == "AA",
        list(mean(arr_delay), mean(dep_delay)),
        by = list(origin, dest, month)]
ans
```
b) Sorted by: keyby
```{r}
ans <- flights[carrier == "AA",
        list(mean(arr_delay), mean(dep_delay)),
        keyby = .(origin, dest, month)]
ans
```
c) Chaining


The code with out chaining.
```{r}
ans <- flights[carrier == "AA",list(.N), by = list(origin, dest)]
ans <- ans[order(origin, -dest)]
head(ans)
```
the code with chaining.
```{r}
ans <- flights[carrier == "AA", .N, by = list(origin, dest)][order(origin, -dest)]
ans
```

d) Expressions in by
```{r}
ans <- flights[, .N, list(dep_delay>0, arr_delay>0)]
ans
```
e) Multiple columns in j - .SD
```{r}
DT
DT[, print(.SD), by = ID]
```

```{r}
DT[, lapply(.SD, mean), by = ID]
```
```{r}
flights[,.SD,by = origin]
```
How can we specify just the columns we would like to compute the mean() on?
```{r}
flights[carrier == "AA",                       ## Only on trips with carrier "AA"
        lapply(.SD, mean),                     ## compute the mean
        by = .(origin, dest, month),           ## for every 'origin,dest,month'
        .SDcols = c("arr_delay", "dep_delay")] ## for just those specified in .SDcols

```
f) Subset .SD for each group:


How can we return the first two rows for each month?
```{r}
ans <- flights[, head(.SD, 2), by = month]
head(ans)
```
g) Why keep j so flexible?

How can we concatenate columns a and b for each group in ID?
```{r}
DT
```

```{r}
DT[, list(val = c(a,b)), by = ID]
```
```{r}
DT
```

```{r}
## (1) look at the difference between
DT[, print(c(a,b)), by = ID]
# [1] 1 2 3 7 8 9
# [1]  4  5 10 11
# [1]  6 12
# Empty data.table (0 rows) of 1 col: ID

## (2) and
DT[, print(list(c(a,b))), by = ID]
# [[1]]
# [1] 1 2 3 7 8 9
# 
# [[1]]
# [1]  4  5 10 11
# 
# [[1]]
# [1]  6 12
# Empty data.table (0 rows) of 1 col: ID
```
setDT is much better than as.data.table ,this difference is memory use,setDT used only for list,data.frame,data.table

the as.data.table is used for any R object.
```{r}
#print(as.data.table(flights[[4]]))
class(print(as.data.table(flights[[4]])))
```
keys concept in data.table.
```{r}
set.seed(1L)
DF = data.frame(ID1 = sample(letters[1:2], 10, TRUE),
                ID2 = sample(1:3, 10, TRUE),
                val = sample(10),
                stringsAsFactors = FALSE,
                row.names = sample(LETTERS[1:10]))
DF
rownames(DF)
DF["C", ]
```
Note that row names have been reset.

data.tables never uses row names. Since data.tables inherit from data.frames, it still has the row names attribute. But it never uses them. We'll see in a moment as to why.

If you would like to preserve the row names, use keep.rownames = TRUE in as.data.table() - this will create a new column called rn and assign row names to this column.
```{r}
DT = as.data.table(DF)
rownames(DT)
```
Instead, in data.tables we set and use keys. Think of a key as supercharged rownames.

Keys and their properties
We can set keys on multiple columns and the column can be of different types - integer, numeric, character, factor, integer64 etc. list and complex types are not supported yet.

Uniqueness is not enforced, i.e., duplicate key values are allowed. Since rows are sorted by key, any duplicates in the key columns will appear consecutively.

Setting a key does two things:

physically reorders the rows of the data.table by the column(s) provided by reference, always in increasing order.

marks those columns as key columns by setting an attribute called sorted to the data.table.

Since the rows are reordered, a data.table can have at most one key because it can not be sorted in more than one way.




concept :-
set, get and use keys on a data.table
note :- we can also set key while creating the data.table object by passing the parameter key in the function.
```{r}
setkey(flights, origin)
head(flights)
```
You can use the function setkey() and provide the column names (without quoting them). This is helpful during interactive use.

Alternatively you can pass a character vector of column names to the function setkeyv(). This is particularly useful while designing functions to pass columns to set key on as function arguments.

Note that we did not have to assign the result back to a variable. This is because like the := function we saw in the "Introduction to data.table" vignette, setkey() and setkeyv() modify the input data.table by reference. They return the result invisibly.

The data.table is now reordered (or sorted) by the column we provided - origin. Since we reorder by reference, we only require additional memory of one column of length equal to the number of rows in the data.table, and is therefore very memory efficient.

You can also set keys directly when creating data.tables using the data.table() function using key argument. It takes a character vector of column names.



- Use the key column origin to subset all rows where the origin airport matches "JFK"
```{r}
flights[list("JFK")]
```
We can subset any amount of values as required
```{r}
flights[c("JFK", "LGA")]   ## same as flights[.(c("JFK", "LGA"))]
```
we can get key value by key function.
```{r}
key(flights)
```
c) Keys and multiple columns

How can I set keys on both origin and dest columns?
```{r}
setkey(flights, origin, dest)
head(flights)
```

Subset all rows using key columns where first key column origin matches "JFK" and second key column dest matches "MIA"
```{r}
flights[list("JFK", "MIA")]
```
Subset all rows where just the first key column origin matches "JFK"
```{r}
key(flights)
flights[.("JFK")]

```

Subset all rows where just the second key column dest matches "MIA"
```{r}
View(flights[.(unique(origin), "MIA")])
```

2) Combining keys with j and by
a) Select in j

```{r}
flights[list("LGA", "TPA"), list(arr_delay)]
```

```{r}
flights[list("LGA", "TPA"), "arr_delay", with = FALSE]
```
b) Chaining

On the result obtained above, use chaining to order the column in decreasing order.
```{r}
flights[list("LGA", "TPA"), list(arr_delay)][order(-arr_delay)]
```

```{r}
flights[list("LGA", "TPA"), max(arr_delay)]
flights1<-flights
```
d) sub-assign by reference using := in j


1. Reference semantics
All the operations we have seen so far in the previous vignette resulted in a new data set. We will see how to add new column(s), update or delete existing column(s) on the original data.
b) The := operator
```{r}
#The LHS := RHS form

DT[, c("colA", "colB", ...) := list(valA, valB, ...)]

# when you have only one column to assign to you
# can drop the quotes and list(), for convenience
DT[, colA := valA]
#The functional form

DT[, `:=`(colA = valA, # valA is assigned to colA
	  colB = valB, # valB is assigned to colB
	  ...
)]
```

2. Add/update/delete columns by reference
a) Add columns by reference
```{r}
flights1[, `:=`(speed = distance / (air_time/60), # speed in mph (mi/h)
	       delay = arr_delay + dep_delay)]   # delay in minutes
head(flights1)
```
b) Update some rows of columns by reference - sub-assign by reference
```{r}
# get all 'hours' in flights
flights1[, sort(unique(hour))]
```
- Replace those rows where hour == 24 with the value 0
```{r}
flights1[hour == 24L, hour := 0L][]
```

```{r}
# check again for '24'
flights1[, sort(unique(hour))]
```
c) Delete column by reference
c) Delete column by reference
- Remove delay column
```{r}
flights1[, c("delay") := NULL]
head(flights1)
## or using the functional form
# flights[, `:=`(delay = NULL)]
```
d) := along with grouping using by
We have already seen the use of i along with := in Section 2b. Let's now see how we can use := along with by.

- How can we add a new column which contains for each orig,dest pair the maximum speed?
```{r}
flights1[, max_speed := max(speed), by = list(origin, dest)]
head(flights1)
```
e) Multiple columns and :=`
```{r}
in_cols  = c("dep_delay", "arr_delay")
out_cols = c("max_dep_delay", "max_arr_delay")
flights[, c(out_cols) := lapply(.SD, max), by = month, .SDcols = in_cols]
head(flights1)
```
```{r}
flights1[, c("speed", "max_speed", "max_dep_delay", "max_arr_delay") := NULL]
head(flights1)
```
3) := and copy()
```{r}
DT = data.table(x = 1L, y = 2L)
DT_n = names(DT)
DT_n
```

```{r}
DT_n = copy(names(DT))
```

```{r}
setkey(flights, hour)
key(flights)
# [1] "hour"
flights[list(24), hour := 0L]
key(flights)
```

```{r}
flights[, sort(unique(hour))]
#  [1]  0  1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 19 20 21 22 23
```
e) Aggregation using by
Let's set the key back to origin, dest first.
```{r}
setkey(flights, origin, dest)
key(flights)
```
- Get the maximum departure delay for each month corresponding to origin = "JFK". Order the result by month
```{r}
ans <- flights["JFK", max(dep_delay), keyby = month]
head(ans)
```
3) Additional arguments - mult and nomatch
a) The mult argument
```{r}
flights[list("JFK", "MIA"), mult = "first"]
```
```{r}
flights[.(c("LGA", "JFK", "EWR"), "XNA"), mult = "last"]
```


```{r}
flights[.(c("LGA", "JFK", "EWR"), "XNA"), mult = "last"]
```
b) The nomatch argument
We can choose if queries that do not match should return NA or be skipped altogether using the nomatch argument.
```{r}
flights[.(c("LGA", "JFK", "EWR"), "XNA"), mult = "last", nomatch = NULL]
```
4) binary search vs vector scans
We have seen so far how we can set and use keys to subset. But what's the advantage? For example, instead of doing:

One advantage very likely is shorter syntax. But even more than that, binary search based subsets are incredibly fast.

As the time goes data.table gets new optimization and currently the latter call is automatically optimized to use binary search.
```{r}
# key by origin,dest columns
flights[list("JFK", "MIA")]
```
1. Secondary indices
a) What are secondary indices?

b) Set and get secondary indices
- How can we set the column origin as a secondary index in the data.table flights?
```{r}
setindex(flights, origin)
head(flights)
```

```{r}
c<-names(attributes(flights))
```

```{r}
indices(flights)
setindex(flights, origin, dest)
indices(flights)
```
c) Why do we need secondary indices?

- Reordering a data.table can be expensive and not always ideal
Consider the case where you would like to perform a fast key based subset on origin column for the value "JFK". We'd do this as:
```{r}
setkey(flights, origin)
flights["JFK"]
```

```{r}
setkey(flights, dest)
flights["LAX"]
```

```{r}

```

